\chapter{Problems on arrays}
Arrays are the simplest data structures.
An array $A$ of length $n$ is a sequence of $n$ elements, $A[0], \dots, A[n - 1]$.
If each element is also an array, the array is called a \emph{matrix}.
An array of $n$ arrays, in which each $A[i]$ is an array of length $m$, is called an $n \times m$ \emph{matrix}.

\paragraph{Notation}
For an array $A$, $A[i \dots j]$ denotes the subarray $A[i] \dots A[j]$.
For two arrays $A$ and $B$, $A + B$ denotes the concatenation of $A$ and $B$.

\section{Find specified element(s)}
In this section, we focus on the problems of finding specified element(s) in a given array or matrix.

\begin{Exercise}
Let $A$ be an array of $n$ integers, in which all integers are between 0 and $n - 2$.
Moreover, all integers between 0 and $n - 2$ appear at least once in $A$; that is, there is only one duplicate integer in $A$.
Design an algorithm to find the duplicate elements in $A$ without modifying the content of $A$.
\end{Exercise}

\begin{Answer}
Since each integer in $A$ is between $0$ and $n - 2$, each element in $A$ can be considered as a node in a linked list; that is, the $i$-th element points to the $A[i]$-th element.
Finding the duplicate integer, $k = A[i] = A[j]$ for some indices $i$ and $j$, is equivalent to finding the $k$-th element, pointed by exactly two elements: the $i$-th element and the $j$-th element.
Since $n - 1$ is not in $A$, the $(n - 1)$-th element can be considered as the head of the linked list. 
Since two elements are pointing to the $k$-th element, a cycle exists in the linked list, and the head of the cycle must be the $k$-th element.
Thus, we can use the \href{https://en.wikipedia.org/wiki/Cycle_detection}{cycle detection algorithm} to find the head of the cycle, which is the duplicate integer.

\subparagraph{Complexity analysis} The time complexity of this algorithm is linear, and the space complexity is $O(1)$.
\begin{remark}
This solution is discussed in Giant's paper~\cite{Ginat2012}.
This problem appears on \href{https://leetcode.com/problems/find-the-duplicate-number/}{LeetCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=Majority]
Given an integer $k > 1$ and an array $A$ of $n$ integers, design an algorithm to find all integers occurring more than $\frac{n}{k}$ times in $A$ if such an integer exists.
\end{Exercise}

\begin{Answer}
The key observation for this problem is as follows: let $c$ be an integer occurring more than $\frac{n}{k}$ times in $A$. If $k$ distinct integers are removed from $A$, then $c$ still occurs more than $\frac{n - k}{k}$ times among the remaining integers.

Based on this observation, we maintain a set $S$ of at most $k - 1$ candidates. For each candidate, we also maintain its frequency so far.
Initially, $S$ is an empty set.
The algorithm reads the integers in $A$ one by one. 
Let $x$ be the integer read by the algorithm.
If $x$ is already in $S$, increase the frequency of $x$ by one. 
If $x$ is not in $S$ and $\abs{S}$ is smaller than $k$, then insert $x$ into $S$ and set $x$'s frequency to one. 
Otherwise, decrement all frequencies of all candidates in $S$ by one; this step is equal to removing $k$ distinct integers from $A$.

After processing all integers in $A$, all integers occurring more than $\frac{n}{k}$ times in $A$ should be in $S$, and each candidate can be verified in linear time.
\subparagraph{Complexity analysis} The time complexity of this algorithm is $O(nk)$.
\begin{remark}
This solution is discussed in Misra's paper~\cite{Misra1982}.
When $k = 2$, this problem is called the \emph{majority finding} problem~\cite{Moore1991}.
For any positive integer $k$, all integers occurring more than $\frac{n}{k}$ times can be found in linear time independent of $k$~\cite{Karp2003}. 
This problem appears on \href{http://www.lintcode.com/en/problem/majority-number-iii/}{LintCode} and \href{https://leetcode.com/problems/majority-element-ii/}{LeetCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=Saddleback search]
Given an integer $x$  and an $n \times m$ matrix $M$ of integers, $m \leq n$, in which all rows and columns are in non-decreasing order, design an algorithm to determine whether $x$ in $M$.
\end{Exercise}
\begin{Answer}
The key observation for this problem is as follows: for any pair of indices , $i$ and $j$, if $M[i][j] > x$, then $M[k][j]$ is larger than $x$ for all $i \leq k < n$. Similarly, if $M[i][j] < x$, then $M[i][k]$ is smaller than $x$ for all $0 \leq k \leq j$.

Based on this observation, the algorithm can eliminate either a row or a column after comparing an entry with $x$.

We always compare $x$ of the top-right element of the remaining submatrix. If this entry equals $x$, then we found the answer. If this entry is smaller $x$, then the top row can be discarded. If this entry is larger than $x$, the the rightmost column can be discarded. This technique is called the \emph{saddleback search}.
\subparagraph{Complexity analysis} Since $M$ has $O(n + m)$ rows and columns, the time complexity is $O(n + m)$. 

\paragraph{A better solution} Let $c = \lfloor m / 2 \rfloor$. First, use binary search to find an index $i$ in the middle column such that $M[i][c] \leq x \leq M[i + 1][c]$. 
If one of $M[i][c]$ and $M[i + 1][c]$ equals $x$, then we found the answer. 
Otherwise, the submatrix in the top-left of $M[i][c]$ cannot contain $x$. Similarly, the submatrix in the bottom-right of $M[i + 1][c]$ cannot contain $x$. We can recursively search for $x$ in the remaining two submatrices. 
\subparagraph{Complexity analysis}  The time complexity is $O(m \lg \frac{n}{m})$.
\begin{remark}
This solution is from Bird's paper~\cite{Bird2006}. This problem appears on \href{https://leetcode.com/problems/search-a-2d-matrix-ii/}{LeetCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=Selection in a sorted matrix]
Given a positive integer $k$ and an $n \times m$ matrix $M$ of integers, in which all rows and columns are in non-decreasing order, design an algorithm to find the $k$-th smallest integer in $M$.
\end{Exercise}

\begin{Answer}
Since the $k$-th smallest integer must be between $min = \min_{i, j} M[i][j]$ and $max = \max_{i, j} M[i][j]$, the $k$-th smallest integer can be found by using the binary searching on the integers $[min, \dots, max]$.

Note that for any integer $x$, we can determine the rank of $x$ in $O(n + m)$ time by using the saddleback search. 
Let $[l ,\dots, h]$ be the range of integers to be searched.
If the rank of $mid = \lfloor (l + h) / 2 \rfloor$ is $k$, then we found the answer.
If the rank of $mid$ is larger than $k$, then the search range can be reduced to $[l, \dots, mid - 1]$.
Otherwise, the search range can be reduced to $[mid + 1, \dots, h]$.

\subparagraph{Complexity analysis} Since the maximum number of iterations is $\lg (max - min)$, the time complexity is $O((n + m) \lg (max - min))$.

This solution has two drawbacks: one minor drawback is that $\lg (max - min)$ is in fact linear in the length of the input. 
Theoretically, this method is not better than applying \href{https://en.wikipedia.org/wiki/Median_of_medians}{linear time selection algorithm} on $M$.

Another drawback is that this method can only be applied on numbers or data like numbers, since this method needs to find the middle of two elements in $M$, even if the middle element is not in $M$. Thus, this method is not general.

\paragraph{A better solution} We use a \href{https://en.wikipedia.org/wiki/Prune_and_search}{prune-and-search} method to iteratively eliminate elements that cannot be the answer until only one element remains.

In each iteration, let $m_i$ be the median of the remaining elements in the $i$-th column.
Let $m^*$ be the median of the set $\{ m_i \mid 0 \leq i < m \}$ and let $r$ be the rank of $m^*$ in $M$.
If $r$ is $k$, then we found the answer.
If $r$ is larger than $k$, then for each column $i$ such that $m_i \geq m^*$, all elements at least $m_i$ can be eliminated.
If $r$ is smaller than $k$, then for each column $i$ such that $m_i \leq m^*$, all elements at most $m_i$ can be eliminated.

\subparagraph{Complexity analysis} In each iteration, $m^*$ can be determined in $O(m)$ time, and the rank of $m^*$ can be computed in $O(m + n)$ time. Since half of columns are halved in each iteration, the number of iterations is $O(\lg n)$. Thus, the time complexity is $O((m+n) \lg n)$.

\begin{remark}
This problem can be solved in $O(n)$ time~\cite{Frederickson1982,Frederickson1984,Mirzaian1985}. This problem appears on \href{http://www.lintcode.com/en/problem/kth-smallest-number-in-sorted-matrix/}{LintCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=Intersection of two sorted arrays]
Given two sorted arrays $A$ and $B$, where $\abs{A} = n$ > m =$\abs{B} = m$ and the elements in individual array are distinct, design an algorithm to find the intersection of these two arrays.
\end{Exercise}

\begin{Answer}
Use an idea similar to merge two arrays, the intersection can be found in $O(n + m)$ time. 
However, when $n \gg m$, this algorithm is slow. 
In the case of $n \gg m$, using binary search to find the occurrence of each element of $B$ in $A$ is more efficient, whose time complexity is $O(m \lg n)$.

\subparagraph{A better solution}
One key observation is as follows:
Let $f(i)$ be the index of $A$ satisfying $A[j] \leq B[i] < A[j + 1]$.
For each index $i$, we know that $B[i']$, $i' < i$, cannot be equal to any element $A[j]$, where $j' > f(i)$.
Moreover, the value of $f(i)$ can be computed in $O(\lg n)$ time by using a binary search.

Thus, we can use the median, $B[i]$, of $B$ to subdivide $B$ into two subarrays, and subdivide $A$ into two subarrays at the location $A[f(i)]$.
Then, recursively solve these two independent problems.

\subparagraph{Complexity analysis} 
The time complexity is $O(m \lg \frac{n}{m})$.

\begin{remark}
This solution is discussed in Baeza-Yates's and Salinger's paper~\cite{Baeza-Yates2010}.
Demaine \textit{et al.} proposed a different approach based on exponential search~\cite{Demaine2000}.
The unsorted version appears on \href{http://www.lintcode.com/en/problem/intersection-of-two-arrays/}{LintCode} and \href{https://leetcode.com/problems/intersection-of-two-arrays/}{LeetCode}.
\end{remark}
\end{Answer}

\subsection{Other problems}
Here are some other problems:
\begin{enumerate}
\item \href{https://en.wikipedia.org/wiki/3SUM}{3-sum}: given an array $A$ of $n$ integers, three elements in $A$ whose sum equals zero can be found in $o(n^2)$ time~\cite{Jorgensen2014}. This problem appears on \href{https://leetcode.com/problems/3sum/}{LeetCode}.
\item \href{https://en.wikipedia.org/wiki/All_nearest_smaller_values}{all nearest smaller values}: given an array $A$ of $n$ integers, an array $S$, in which $S[i]$ is the largest index $j < i$ subject to $A[j] < A[i]$, can be computed in linear time. That is, for all $0 \leq i < n$, $S[i] = \max \{ j \mid j < i \text{ and } A[j] < A[i] \}$ can be computed in linear time. The solution of this problem can be used to solve the largest rectangle in histogram problem on \href{https://leetcode.com/problems/largest-rectangle-in-histogram/}{LeetCode}. The solution of the largest rectangle in histogram problem can be used to solve the \href{http://www.drdobbs.com/database/the-maximal-rectangle-problem/184410529}{maximal rectangle problem} on \href{https://leetcode.com/problems/maximal-rectangle/}{Leetcode}.
\item inversion counting: given an array $A$ of $n$ integers, the number of \href{https://en.wikipedia.org/wiki/Inversion_(discrete_mathematics)}{inversions} can be computed in $O(n \lg n)$ time. This problem appears on \href{https://leetcode.com/problems/count-of-smaller-numbers-after-self/}{LeetCode}.
\end{enumerate}

\section{Find subarray(s) or subsequence(s)}
In this section, we focus on the problems of finding specified subarray(s) or subsequences in a given array.

\begin{Exercise}[title=Path partition]
Given a positive integer $k$ and an array $A$ of $n$ integers, design an algorithm to partition $A$ into $k$ subarrays so that the maximum subarray sum among these $k$ subarrays is minimized.\\
That is, find a sequence of $k$ pairs of indices $\langle (b_1, e_1) \cdots (b_k, e_k) \rangle$ minimizing $\max_{1 \leq i \leq k} \sum_{j = b_i}^{e_i} A[j]$ subject to
\begin{enumerate} 
\item $b_1 = 0$ and $e_k = n - 1$,
\item $b_i \leq e_i$ for all $1 \leq i \leq k$,
\item and $b_{i + 1} = e_i + 1$ for all $1 \leq i < k$.
\end{enumerate}
\end{Exercise}
\begin{Answer}
For a number $d$, whether these $n$ integers can be partitioned into $k$ parts with the maximum subarray sum at most $d$ can be verified in linear time by using a greedy approach.
Let $S = \sum_{i=0}^{n-1} A[i]$ and let $d^*$ be the minimum integer such that $A$ can be partitioned into $k$ parts with the maximum subarray sum at most $d^*$. 
Since $d^*$ must be between $\lceil \frac{S}{k} \rceil$ and $S$, $d^*$ can be found by using a binary search on integers between $\lceil \frac{S}{k} \rceil$ and $S$.
\subparagraph{Complexity analysis} The time complexity is $O(n \lg \frac{S(k-1)}{k})$.

\paragraph{A better solution}  Note that $d^*$ must be $\sum_{k=i}^j A[k]$ for some $i$ and $j$.
Create a matrix $M$, such that $M[i][j] = \sum_{k=i}^j A[k]$, in which all rows and columns in $M$ are in increasing order.
Finding a pair of indices, $i$ and $j$, such that $M[i][j]$ is $d^*$, is our goal.

We use a \href{https://en.wikipedia.org/wiki/Prune_and_search}{prune-and-search} method to iteratively eliminate elements that cannot be the answer until only one element remains.
Since creating $M$ takes $O(n^2)$ time, the algorithm should not explicitly create $M$.
Instead, the algorithm maintains a lower bound and an upper bound for each column to represent the current remaining elements in each column.

In each iteration, let $m_i$ be the median of the remaining elements in the $i$-th column.
Let $m^*$ be the median of the set $C = \{ m_i \mid 0 \leq i < m \}$.
If $A$ can be partitioned into $k$ parts with the maximum subarray sum at most $m^*$, then $d^*$ is at most $m^*$.
Otherwise, $d^*$ should be greater than $m^*$.
All candidates that cannot be the answer can be pruned by using a saddle back search.

\subparagraph{Complexity analysis} In each iteration, $m^*$ can be determined in $O(n)$ time. 
Since the saddleback search also takes $O(n)$ time, each iteration takes $O(n)$ time as well. 
Because at least a quarter of candidates are eliminated in each iteration, the time complexity is $O(n \lg n)$.
\begin{remark}
This solution is from the paper of Khanna \textit{et al.}~\cite{Khanna1997}. This problem can be solved in linear time~\cite{Frederickson1991}. This problem appears on \href{http://www.lintcode.com/en/problem/copy-books/}{LintCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=$k$-covers]
Given a positive integer $k$ and an array $A$ of $n$ non-zero integers, design an algorithm to find $k$ disjoints and non-empty subarrays, called a $k$-cover, with the maximum sum.
\end{Exercise}
\begin{Answer}
Without loss of generality, all consecutive positive/negative integers can be merged into one positive/negative integer. 
Thus, we assume that $A$ has integers with alternating signs, and the integers at the both ends are positive; that is $\sign(A[i]) \neq \sign(A[i+1])$ for all $0 \leq i < n - 1$, and both $A[0]$ and $A[n - 1]$ are positive, where $n \geq 2k - 1$ is an odd integer.

Creating a prefix sum array $P$, $P[i] = \sum_{j = 0}^i A[i]$, for all $0 \leq i < n$ is a common trick for solving problems involving subarray sum, since the sum of $A[b\dots e] = \sum_{i = b}^{e - 1} A[q]$ becomes $P[e] - P[b - 1]$ assuming $P[-1] = 0$.
The problem then becomes 
\begin{mdframed}
\centering
find a sequence of disjoint index intervals $\langle [b_0, e_0], \dots, [b_{k-1}, e_{k-1}]\rangle$ maximizing $\sum_{i = 0}^{k - 1} P[e_i] - P[b_i]$.
\end{mdframed}

Since the signs of integers alternate in $A$, $P[i - 1]$ is smaller than  $P[i]$ for all even $0 \leq i \leq n - 1$ and $P[i]$ is larger than  $P[i + 1]$ for all even $0 \leq i < n - 1$.
Thus, in an optimal solution, each interval begins at an odd index and ends at an even index.
For convenience, the \emph{value} of an index interval $[b, e]$ is defined as $P[e] - P[b]$, and its \emph{value interval} is defined as $[P[b], P[e]]$.
An index interval $[b, e]$ with a positive value satisfying $b$ is odd and $e$ is even is called a \emph{useful} interval.
Any optimal solution only contains useful intervals.

Now, consider two useful intervals $[b, e]$ and $[b', e']$, where $e < b'$.
For these two intervals, four possibilities exist in an optimal solution.
An optimal solution may take none of these intervals, may take both intervals, may take exactly one of the intervals, or may take the interval $[b, e']$.
We can rule out some possibilities based on the relative order of the value intervals of $[b, e]$ and $[b', e']$.

Let $[l, h]$ and $[l', h']$ be the value intervals of $[b, e]$ and $[b', e']$ respectively.
If $l \geq l'$, for any solution containing $[b, e'']$ where $e' \leq e''$, another solution can be constructed by replacing the interval $[b, e'']$ by $[b', e'']$, and the resulting solution is no worse than the original solution. 
Thus, combining the interval $[b, e]$ with other intervals after $[b', e']$ is not necessary.

Based on this observation, we maintain a stack $S$ of value intervals so that
\begin{enumerate}
\item let $[l_0, h_0] \dots [l_t, h_t]$ be their value intervals from the bottom to the top. The value interval $[l_i, h_i]$ strictly includes $[l_{i + 1}, h_{i + 1}]$ for all $0 \leq i < t - 1$.
\item let $[b_0, e_0] \dots [b_t, e_t]$ be their corresponding index intervals from the bottom to the top. These intervals are non-empty and satisfy $e_i < b_{i + 1}$ for all $0 \leq i < t - 1$.
\end{enumerate}

All useful index intervals having the following form $[b, b + 1]$ are processed in order.
Let $[l, h]$ be the left end of the top value interval in the stack.
For each value interval $[l', h']$, three cases are possible:
\begin{enumerate}
\item $l \geq l'$: since the top value interval will never be combined with any interval after the current interval, the top interval is popped from $S$, and the value of the top interval, $h_t - l_t$, is stored in a list $L$.
\item $l < l'$ and $h > h'$: push the current interval to $S$.
\item $l < l'$ and $h \leq h'$: merge the top interval into the current interval. That is, the top interval is popped, and change the current interval to $[l, h']$. Since an optimal solution may need to use these two intervals separately if $h - l' > 0$, an additional value $h - l'$ is stored in $L$.
\end{enumerate}
In the end, the $k$ largest values in $L$ and $S$ are selected, which will be the maximum sum of $k$ disjoint subarrays.

\subparagraph{Complexity analysis} Since each index interval can be pushed and popped at most once, the time complexity is $O(n)$.
\begin{remark}
%This problem was firstly considered by Cs{\"{u}}r{\"{o}}s~\cite{Csuros04}.
Bengtsson and Chen designed the first linear time algorithm for this problem~\cite{Gawrychowski2015}.
%Gawrychowski and Nicholson designed a data structure that can be built in linear time and supports $k$-cover query $O(k)$ time~\cite{Gawrychowski2015}.
This problem appears on \href{http://www.lintcode.com/en/problem/maximum-subarray-iii/}{LintCode} and \href{https://leetcode.com/problems/best-time-to-buy-and-sell-stock-iv/}{LeetCode}.
When $k = 1$, the problem can be solved by using \href{https://en.wikipedia.org/wiki/Maximum_subarray_problem#Kadane.27s_algorithm}{Kadane's algorithm}. This special case appears on \href{https://leetcode.com/problems/maximum-subarray/}{LeetCode}.
\end{remark}
\end{Answer}

\subsection{Other problems}
Here are some other problems:
\begin{enumerate}
\item \href{https://en.wikipedia.org/wiki/Longest_increasing_subsequence}{longest increasing subsequence}: this problem can be solved in $n \lg n + O(n)$ time~\cite{Fredman1975}. This problem appears on \href{https://leetcode.com/problems/longest-increasing-subsequence/}{LeetCode}.
\end{enumerate}

\section{Manipulate}
In this section, we focus on the problems of manipulating a given array.

\begin{Exercise}
Given a positive integer $k$ and an array $A$ of $n$ integers, design an algorithm to modify $A$ into another array $B$ of $n$ integers in which $\abs{B[i] - B[i + 1]} \leq k$ for all $0 \leq i < n - 1$, while minimizing $\sum_{i=0}^{n - 1} \abs{A[i] - B[i]}$.
\end{Exercise}
\begin{Answer}
This problem can be formulated as a linear programming problem.
Two set of variables, $X$ and $Y$, are used to represent the result $B[i] = A[i] + X[i] - Y[i]$. 
The objective function is the summation of $X$ and $Y$. 
By the difference constraint, $ -k \leq B[i] - B[i+1] \leq k$ is required, and this can be formulated as $-k \leq  A[i] + X[i] - Y[i] - A[i+1] - X[i+1] + Y[i+1] \leq k$. 
After algebraic manipulation and transform into standard form by introducing slack variables $S$ and $T$, the following two equality constraints can be obtained: 
\[ X[i] - Y[i] - X[i+1] + Y[i+1] + S[i] = k - P[i] \text{ and } X[i] - Y[i] - X[i+1] + Y[i+1] - T[i] = -k - P[i],\]
 where $P[i] = A[i] - A[i+1]$. 

From these two equality constraints, $S[i] + T[i]$ must be equal to  $2k$. 
Thus, one of the equality constraint is removed, and an upper bound $2k$ to $S[i]$ or $T[i]$ is introduced.
If one one redundant constraint by summing all $n - 1$ equality constraints is introduced and all constraints are negated on both sides, the formulation becomes a min-cost circulation problem.
Since the underlying graph is a series-parallel graph and min-cost circulation on series-parallel graphs can be solved in $O(n \lg n)$ time, this problem can be solved in $O(n \lg n)$ time as well.
\begin{remark}
This solution is from \href{http://www.chaoxuprime.com/posts/2015-01-27-bounded-regression-on-data-streams.html}{here}.
Booth and Tarjan designed an $O(n \lg n)$-time algorithm for finding a min-cost circulation on series-parallel graphs~\cite{Booth1993}.
This problem appears on \href{http://www.lintcode.com/en/problem/minimum-adjustment-cost/}{LintCode}.
\end{remark}
\end{Answer}

\begin{Exercise}[title=Block interchange]
Given two indices $a$ and $b$ and an array $A$ of $n$ integers, design an algorithm to interchange $A[0 \dots a]$ with $A[b \dots n - 1]$.
That is, initially $A$ equals $P + M + S$, where $P = A[0 \dots a]$, $M = A[a+1 \dots b-1]$, and $S = A[b \dots n - 1]$, but $A$ equals $S + M + P$ in the end.
\end{Exercise}
\begin{Answer}
One easy solution is as follows:
First, reverse the whole string so that the string becomes $A^R = S^R + M^R + P^R$.
Then, reverse $S^R$, $M^R$, and $P^R$ respectively.
However, this algorithm is not optimal.
For example, when $\abs{P} = \abs{S}$, only $\abs{P}$ swaps are required, but the algorithm uses $n$ swaps.

\paragraph{A better solution}  The idea is to move characters to the desired locations directly.
If $\abs{P} \leq \abs{S}$, then first swap $P$ with the last $\abs{P}$ characters of $S$.
The intermediate result is $S'' + M + S' + P$, where $S = S' + S''$ and $\abs{S''} = \abs{P}$. 
Since the desired result is $S + M + P$, the algorithm recursively interchanges $S'' + M$ with $S'$ in the subarray $S'' + M + S'$.

If $\abs{P} > \abs{S}$, then first swap the first $\abs{S}$ characters of $P$ with $S$.
The intermediate result is $S + P'' + M + P'$, where $P = P' + P''$ and $\abs{P'} = \abs{S}$. 
Since the desired result is $S + M + P$, the algorithm recursively interchanges $P''$ with $M + P'$ in the subarray $P'' + M + P'$.

\subparagraph{Complexity analysis} The algorithms uses $n - \text{gcd}(\abs{P} + \abs{M}, \abs{M} + \abs{S})$ swaps.
\begin{remark}
This solution is from Mohammed and Subi's paper~\cite{Mohammed1987}. 
For the special case of rotating an array circularly, Bentley proposed a juggling algorithm~\cite{Bentley1999}.
Shene analyzed the time complexity and the lower bound for several algorithms~\cite{Shene1997}.
Benchmark of these algorithms can be found on~\href{http://www.drdobbs.com/parallel/benchmarking-block-swapping-algorithms/232900395}{Dr.Dobb's}.
The fastest algorithm is Gries-Mills algorithm~\cite{Gries1981}.
The problem of rotating an array circularly appears on \href{https://leetcode.com/problems/rotate-array/}{LeetCode}.
\end{remark}
\end{Answer}

\subsection{Other problems}
Here are some other problems:
\begin{enumerate}
\item matching nuts and bolts: given $n$ nuts and $n$ bolts, every bolt matches exactly one nuts. You can try to match a nut and a bolt and see which one is bigger or they can be matched. 
However, you cannot compare two bolts or two nuts. All bolts and nuts can be matched in $O(n \lg n)$ time deterministically~\cite{Komlos1998}. This problem can also be solved easily by using the randomized quick sort.\footnote{Using Lomuto partition is easier.}  This problem appears on \href{http://www.lintcode.com/en/problem/nuts-bolts-problem/}{LintCode}.
\item partition: given an integer $k$ and an array $A$ of $n$ integers, $A$ can be rearranged in linear time so that all integers smaller than $k$ are in front of all integers at least $k$ by using \href{https://en.wikipedia.org/wiki/Quicksort#Hoare_partition_scheme}{Hoare's partition}, which is more efficient than \href{https://en.wikipedia.org/wiki/Quicksort#Lomuto_partition_scheme}{Lomuto partition}.\footnote{\url{http://cs.stackexchange.com/questions/11458/quicksort-partitioning-hoare-vs-lomuto}.} This problem appears on \href{http://www.lintcode.com/en/problem/partition-array/}{Lintcode}.
\item \href{https://en.wikipedia.org/wiki/Dutch_national_flag_problem}{Dutch national flag problem}: this problem appears on \href{https://leetcode.com/problems/sort-colors/}{LeetCode}.
\end{enumerate}

\section{Range query}
In this section, we focus on the problems of building data structures to support range queries.
\begin{enumerate}
\item range sum query: given an array $A$ of $n$ integers, build a data structure to represent $A$ supporting the following two queries: 
\begin{enumerate}
\item given a pair of indices $i < j$, compute the subarray sum of $A[i \dots j]$.
\item given an index $i$ and a value $v$, set $A[i]$ to $v$.
\end{enumerate}
This problem can be solved by using a \href{https://en.wikipedia.org/wiki/Fenwick_tree}{Fenwick tree}. This problem appears on \href{https://leetcode.com/problems/range-sum-query-mutable/}{LeetCode}.
\item \href{https://en.wikipedia.org/wiki/Range_minimum_query}{range minimum query problem}: this problem appears on \href{http://www.lintcode.com/en/problem/interval-minimum-number/}{LintCode}.
\end{enumerate}



\printbibliography[heading=subbibliography]
